### 2019-12-03 paper/submission.tex
Working through paper, up to results looks pretty good

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
%\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}

%-------------------------------------------------------------------------
\title{Deep Learning for 3D Brain Activity Scans}

\author{Anonymized for Peer Review\\
University of Massachusetts Amherst\\
140 Governors Dr, Amherst, MA 01002\\
{\tt\small @umass.edu}
}

\maketitle
%\thispagestyle{empty}

% ----------------------------------------------------------------------
% ABSTRACT
%-------------------------------------------------------------------------
\begin{abstract}
 Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
 typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
 and limit the ability to stitch together patterns across a third dimension.

 Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
 imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
 overlooked patterns of activity.

 Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
 image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
 restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
 to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
 what the network can teach us from its findings.

\end{abstract}

% ----------------------------------------------------------------------
% INTRODUCTION
% ----------------------------------------------------------------------
\section{Introduction}\label{sec:introduction}
Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
for mental health research and disease diagnosis.
High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
dimensions, are a lot to manage by hand.

Classification of activity and structures should come quite naturally with deep learning practices.
Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
yet often only as a proof of concept, and limited to two dimensions.
Others are working now to try and find universal architectures for most
neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
application with satisfactory performances across the board.

Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
they identify in classification could surely be adopted to highlight what patterns of activity have informed their
own decision making and perhaps clue researchers in to overlooked details.

Here we'll discuss creating highly accurate classifiers using both 2d and 3d convolutional networks
to learn in the typical manner low level filters, here specifically created for brain scans of spatial activity, using
no pretraining on unrelated datasets and no neuro-imaging feature analysis or reduction~\cite{shi2018feature},
letting the model learn for itself what is important so that it can teach the observer instead.

Following this we'll discuss what this may mean for future works in the highly valuable ability to adapt the classifiers
to the new, related task of unsupervised segmentation in 3 dimensions~\cite{shu2016unsupervised} to mirror what has been
a fruitful field of 2d identification of tumors~\cite{akkus2017deep} for this specific application of telling brain
activity patterns.

% ----------------------------------------------------------------------
% BACKGROUND / RELATED
% ----------------------------------------------------------------------
\section{Background}\label{sec:background}

The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
Netherlands circa 2013 who were presented with various grids of images featuring
either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
have formed restraint with given their success in maintaining healthy body mass indices.
The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
balanced against the hypothesized activation of restraints.

In neuro-imaging, this task is known as passive image viewing.

Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
This combination makes this an excellent selection for a proof of concept application of deep learning as the chances
of success on a less subtle activation pattern, observable to trained researchers, are quite high.
This specific application then can serve as a jumping off point for future works where the activity patterns are less
distinguishable, and may be an interesting addition to the works in brain-pattern controlled prosthetics.
This dataset is also ideal as it comes with a fairly usable ground-truth labeling mechanism, as I am not currently
investigating the capacity for unsupervised differentiation.

The dataset can be found from the excellent
OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}

The dataset follows the BIDS (Brain Imaging Data Structure) layout adopted by many neuroscience researchers.
It consists of a survey detailing the relationship of each of the 30 subjects with food and diet taken after the experimental
proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w Modality fMRI scan seen in figure~\ref{fig:ortho-t1w},
and time series of 370-375 BOLD Modality fMRI scans seen in figure~\ref{fig:ortho-bold}, encompassing approximately 10 minutes
capturing brain activity of the subjects during their exposure to the experimental triggers.

The bold scans are quite sizeable as far as datasets go, given the number of each per subject,
however they are a relatively low resolution, consisting
of 64 sagital by 64 coronal by 30 axial scans, with only a single image channel.

The activation is measured in a positive magnitude visible in figure~\ref{fig:ex_stat_map}, through a helper method in a neuro-imaging specialized
image handling library ~\cite{brett_matthew_2019_3544468}.

I had to manually infer pairings of the scans with the experimental phase for classification training given the events log associated
with each subject and the advertised scan times from the paper.
While the data does appear healthy, there may be some discrepancies of interpretation which would warrant future investigation if the
performance on this particular experiment was the primary focus.

You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.

You may find this project repository here: (personal github link removed for anonymity)
Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.

\begin{figure}
  \includegraphics[width=\linewidth]{images/orthoview_bold.png}
  \caption{Bold Modality Low-Res 3d Experimental Scans of Subject Brain}
  \label{fig:ortho-bold}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{images/example_stat_map.png}
  \caption{Activity Measures from Bold Scans}
  \label{fig:ex_stat_map}
\end{figure}

% ----------------------------------------------------------------------
% APPROACH
% ----------------------------------------------------------------------
\section{Approach}\label{sec:approach}

Beginning with SKLearn~\cite{scikit-learn} data readiness was proven with moderately successful classification using
a simple SVM and then the SKLearn Multi-Layer-Perceptron neural network, getting 70-80\% accuracy, sufficing the
initial sanity check that deep learning would be ready to work on.

% TODO these are relevant citations, right?
Leveraging PyTorch~\cite{paszke2017automatic} first a framework for learning was likewise vetted using a simple fully
connected shallow network, graduating to the state of the art convolutional networks, chiefly focussed
on the ResNet~\cite{he2016deep} architecture styles, which have been developed for video classification~\cite{tran2018closer}
utilizing 3 dimensional convolutions~\cite{hara3dcnns} much in the way 2 dimensional convolutions operate, which have
been shown to hold a slight edge on 2 dimensions~\cite{payan2015predicting} in other applications.

\subsection{Data Preparation}\label{subsec:data-preparation}

First as the data was not prepared for classification, critical cleaning include subtracting the mean image to better
highlight meaningful variance, and standardizing the data between 0 and 1.

Given the nature of the input images being 3d brain scans it's intuitive that examining the brain in 3 dimensions
allows for much more interesting patterns of brain activity and structuring being recognized.

This comes with a substantial cost however, as the extra dimension explodes the memory demands of the input data, and
the number of trainable parameters for the models.
Many of the recommendations following will vary based on the particular hardware, but with a 4GB GPU I had to be quite
careful about batch-sizes to not over-allocate memory.
To combat this, first, a relatively conservative amount of blank space is cropped from the images, halving the total footprint.
Following this, in order to use a batch-size of 64, I interpolate images to approximately $\frac{5}{6}$ their original dimension

In addition, I typically do not use the full cohort of 30 subjects, as I even encounter some RAM restrictions with the
full raw dataset consisting of
$64 sagital \times 64 coronal \times 30 axial \times 32bit floats \times 375 scans \times 30 subjects$ or roughly 5.53 gigabytes

In working with 2 dimensions, omitting the axial dimension for a flat cross section, you'll see the dataset is
$\frac{1}{30}$ the size and hence, much more manageable.

\subsection{Examining the Data}\label{subsec:examining-data}

Two fully processed, correctly classified brain scans are presented in the figure~\ref{fig:comparison}
which pose a daunting challenge to human labelers.

\begin{figure}
  \includegraphics[width=\linewidth]{images/nf_f_hard_comparison.png}
  \caption{Nonfood-Stimuli \& Food-Stimuli Respectively from the same Subject}
  \label{fig:comparison}
\end{figure}

\subsection{Model Selection}\label{subsec:model-selection}

For the primary focus of comparing and improving classification accuracy using 2d and 3d architectures, ResNet models
of the standard 18 layers to 100+ layers are examined, as the ResNet family is both highly competitive, and demonstrated
in both 2 and 3 dimensions which will help comparisons in exploring the benefits of 3d brain scan usage.
Video processing typically limits the layers to a more shallow level than that of image processing, as unfortunately as
with the data, the model parameters explode out of control pretty fast.
The following parameter counts are from standard ResNet architectures adapted to a single input channel.

\begin{center}
 \begin{tabular}{||c c c||}
 \hline
 - & ResNet-18 2d & ResNet-18 3d \\ [0.5ex]
 \hline\hline
 Counts & 11,171,266 & 33,148,482 \\ [1ex]
 \hline
 Ratio & 1.0 & 2.97 \\ [1ex]
 \hline
\end{tabular}
\end{center}

\subsection{Comparisons}\label{subsec:comparisons}

Model performance is typically evaluated using accuracy, precision, recall, and AUC metrics, as well as total time to
train.

% ----------------------------------------------------------------------
% RESULTS
% ----------------------------------------------------------------------
\section{Results}\label{sec:results}
At the date of the milestone, I have shown that my experimental set-up is complete to begin
iterating towards the novel contributions aimed for by this project.

I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
network for classification on my chosen dataset incorporating my own contributions to the process.

I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
excited nevertheless to try.

\subsection{Data}\label{subsec:data}
I have become familiar with the chosen dataset, and created a machine learning friendly format
based off of the data.

I have inferred classification labels from the experimental events timings and the estimated time
of each complete scan in the dataset.

\subsection{Model Architecture}\label{subsec:model-arch}

the project explored minimally the hyperparameters of the chosen model architectures
chiefly was finding the right learning rate for the given normalized data to the model
typically 1e-2 to 1e-4 worked, with 1e-3 being good, and decay being very very helpful
model quite sensitive to adjustments

- scheduler / hyperparams / diff architectures (just go with res3d, res2d+1d, mix three types

\subsection{Scope of Number of Subjects}\label{subsec:num-subjects}

inclusion of more ppts/subjects means more training time (val/test are of same patients,
hence more individual brain shapes means more general, but longer to learn)
20 epochs to see liftoff of 55% with 10 ppts, vs quicker with 6ppts

 \begin{figure}
  \includegraphics[width=\linewidth]{images/2d_cold.png}
  \caption{Cold-Start 2D Training Progress}
  \label{fig:2d_cold}
\end{figure}

\subsection{Warm Starts}\label{subsec:warm-starts}
saving model parameter dict, loading and training again
wowee good stuff, hope to use a LR scheduler in the future to try to control this better?
surprisingly bad results on warm starting expansion to new subjects - old same subjects takes a little while to stabilize
but it gets back to 83% without much fuss, probably LR is a little high to start with and the decay helps a lot.
bounces down for a while, bounces up for a while. breaking 90% which is very cool, might mean warm start has best potential

 \begin{figure}
  \includegraphics[width=\linewidth]{images/2d_warm.png}
  \caption{Warm-Start 2D Training Progress}
  \label{fig:2d_warm}
\end{figure}

\subsection{3 Dimensions}\label{subsec:3-dimensions}
- 3d CNN - typically meant for video classification using time as the third axis,
here we are not using time, though that is a dimension we have access to,
though doesn't make sense for this experiment/project to try to add in
- https://pytorch.org/docs/stable/torchvision/models.html#video-classification

okay it works, and has potential, certainly given the data is richer to be the best, but wow it's frustratingly slow
findings: 3d is cool, but crazy prohibitively slow and not much better than 2d without more refinements. IMHO theres
promise still in exploring it, but given the inexcusable runtime and datasize 2d is probably worth more investment
unless better hardware or serious effort is expended in getting it differentiable

 \begin{figure}
  \includegraphics[width=\linewidth]{images/3d_cold.png}
  \caption{Cold-Start 3D Training Progress}
  \label{fig:3d_cold}
\end{figure}
wrestling with memory constraints (batch size / 3d data huge for cpu/gpu, dataset pretty large in general for ram) slow learning, tapering off learning
different styles of 1d (unfurl 3d, 2d) 2d (mean flattened, one slice) 3d (scale, orig)
LR explorations

future works
better gpu / computer might be important, the 3d models are real fucking beasts to train, and their data is large as fuck
model ensemble (free extra few percent perf)

"none of my alterations to the model architecture worked that well, most did nothing, a few broke things, could still poke around more though"
graphs from 2d training, 3d takes legit 2-3 hours for 200 epochs on 6 subjects with 5./6 & 7. / 2.
batch_size and interpolation hard to juggle too?
3d feels like "luck" plays a big part of getting training to catch/start
seems like scaling down is pretty important and I have better luck with larger batch sizes (meaning scale down is crucial) wonder if aggressive scale down is worthwhile, like over 3./4, image gets real grainy real fast)

cut batch size to avoid scaling down images (TRIED NO BUENO?)

\subsection{Performance}\label{subsec:performance}

 % P-R-F1-S
\begin{center}
 \begin{tabular}{||c c c||}
 \hline
 - & Nonfood & Food \\ [0.5ex]
 \hline\hline
Precision & 0.949074 & 0.958904 \\
 \hline
Recall & 0.957944 & 0.950226 \\
 \hline
F1 & 0.953488 & 0.954545 \\
 \hline
Support & 214 & 221 \\ [1ex]
 \hline
\end{tabular}

 % ROC-AUC / ACC
\end{center}
\begin{center}
 \begin{tabular}{||c c||}
 \hline
 ROC-AUC & Accuracy \\ [0.5ex]
 \hline\hline
  0.9847 & 0.954 \\ [1ex]
 \hline
\end{tabular}
\end{center}

Can get a little higher but these are already encouraging numbers (96 seen) often the model starts leaning a little
one way or the other and this was one of the most balanced runs I've seen

I'm a little unclear about the reliability of my data labeling, so I'm happy with this result, I'd be stoked
to reproduce on 3d, and trusting the model this much, open the door to segmentation

% ----------------------------------------------------------------------
% DISCUSSION
% ----------------------------------------------------------------------
\section{Discussion}\label{sec:discussion}

\subsection{Learning}\label{subsec:learning}
very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at

\subsection{Future Works}\label{subsec:future-works}
Unsupervised segmentation, especially in 3d would be cool
Future works
- could strive to refine 3d more, with more time and compute
- unsupervised segmentation would be cool
- find ways to not have to smoosh images for memory?
- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too

began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
- check out filter viz
- see about unsupervised
- try 3d alternative archs? (dont really need to)

might want to improve 3d before going on to segmentation,

 \begin{figure}
  \includegraphics[width=\linewidth]{images/saliency.png}
  \caption{caption}
  \label{fig:saliency}
\end{figure}

% ----------------------------------------------------------------------
% END
% ----------------------------------------------------------------------
{\small
\bibliographystyle{ieee}
\bibliography{project}
}

% ----------------------------------------------------------------------
% SUPPLEMENTARY
% ----------------------------------------------------------------------
\section{Supplementary Material}\label{sec:supplementary}

Extra material for reference

 \begin{figure}
  \includegraphics[width=\linewidth]{images/orthoview_t1w.png}
  \caption{caption}
  \label{fig:ortho-t1w}
\end{figure}

 \begin{figure}
  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
  \caption{caption}
  \label{fig:cortex-var}
\end{figure}

\end{document}