### 2019-12-03 paper/sections/discussion.tex

\usepackage{graphicx}

\section{Discussion}\label{sec:discussion}

\subsection{Learning}\label{subsec:learning}
very cool! I tried spot checking different scans to see if I could tell the difference visually and I really really
cannot. I think it's encouraging that the model is able to notice impossible to visualize differences and run with it
this well. It's one thing to differentiate a st. bernard from a siamese cat, but two brain activations I'm not good at

\subsection{Future Works}\label{subsec:future-works}
Unsupervised segmentation, especially in 3d would be cool
Future works
- could strive to refine 3d more, with more time and compute
- unsupervised segmentation would be cool
- find ways to not have to smoosh images for memory?
- more visualizations? lack of color kinda sucks, saliency is weird, filters prob too

began work on visualization, it didn't lead anywhere useful, 1 channel might be a problem?
- check out filter viz
- see about unsupervised
- try 3d alternative archs? (dont really need to)

might want to improve 3d before going on to segmentation,

 \begin{figure}
  \includegraphics[width=\linewidth]{images/saliency.png}
  \caption{caption}
  \label{fig:saliency}
\end{figure}