### 2019-12-03 paper/submission.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
%\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[breaklinks=true,bookmarks=false]{hyperref}

\cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}

%-------------------------------------------------------------------------
\title{Deep Learning for 3D Brain Activity Scans}

\author{Anonymized for Peer Review\\
University of Massachusetts Amherst\\
140 Governors Dr, Amherst, MA 01002\\
{\tt\small @umass.edu}
}

\maketitle
%\thispagestyle{empty}

% ------------------------------------------------------------------------
% ABSTRACT
%-------------------------------------------------------------------------
\begin{abstract}
 Analyzing the brain's activity patterns starts with identifying known patterns of well established structures,
 typically observed by human oversight in two dimensions, which may miss subtle, barely visible but useful indicators,
 and limit the ability to stitch together patterns across a third dimension.

 Deep learning's ability to decipher and learn from such nearly impossible patterns in massive datasets,
 imperceptible to preliminary observation makes it an excellent candidate to explore and teach us about
 overlooked patterns of activity.

 Here I showcase convolutional network's ability to correctly classify uncleaned activity data from a passive
 image viewing experiment aimed at triggering activity in particular regions of the brain controlling reward and
 restraint, using the data in both 2d and 3d to explore the merits of adapting video-processing techniques meant
 to apply across time to a third spatial dimension and discuss adapting this to unsupervised segmentation to see
 what the network can teach us from its findings.

\end{abstract}

% ------------------------------------------------------------------------
% BODY
%-------------------------------------------------------------------------
\input{sections/intro.tex}
\input{sections/background.tex}
\input{sections/approach.tex}
\input{sections/results}
\input{sections/discussion.tex}

% ----------------------------------------------------------------------
% END
% ----------------------------------------------------------------------
{\small
\bibliographystyle{ieee}
\bibliography{project}
}

% ----------------------------------------------------------------------
% SUPPLEMENTARY
% ----------------------------------------------------------------------
\section{Supplementary Material}\label{sec:supplementary}

Extra material for reference

 \begin{figure}
  \includegraphics[width=\linewidth]{images/var_on_cortex_2.png}
  \caption{caption}
  \label{fig:cortex-var}
\end{figure}

\end{document}